```{r}
library(biomaRt) #Ensemble Gene 94
ensembl <- useMart(host='http://oct2018.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL')
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)

attributes = c("ensembl_gene_id","external_gene_name","mmulatta_homolog_ensembl_gene","mmulatta_homolog_associated_gene_name","mmulatta_homolog_perc_id","mmulatta_homolog_perc_id_r1","mmulatta_homolog_goc_score","mmulatta_homolog_wga_coverage","mmulatta_homolog_dn","mmulatta_homolog_ds","mmulatta_homolog_orthology_confidence","mmulatta_homolog_orthology_type")

#Get all Human-Macaque gene pairs
humMac <- getBM(attributes = attributes,mart = ensembl)
dim(humMac)

#Get all Human-Macaque gene pairs
hum_type <- getBM(attributes = c("ensembl_gene_id","external_gene_name","gene_biotype"),mart = ensembl)
```

###. Human-Mouse
```{r}
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name", "mmusculus_homolog_perc_id","mmusculus_homolog_perc_id_r1","mmusculus_homolog_goc_score","mmusculus_homolog_wga_coverage","mmusculus_homolog_dn","mmusculus_homolog_ds","mmusculus_homolog_orthology_confidence","mmusculus_homolog_orthology_type")

humMou <- getBM(attributes = attributes,mart = ensembl)
dim(humMou)
```

```{r}
humgene <- unique(humMac$ensembl_gene_id)
humMacMou.AllGene <- merge(humMac, humMou, by = "ensembl_gene_id")
rownames(hum_type) = hum_type$ensembl_gene_id
hum.spe <- humMacMou.AllGene[humMacMou.AllGene$mmulatta_homolog_ensembl_gene=="" & humMacMou.AllGene$mmusculus_homolog_ensembl_gene=="",]
hum.spe.coding <- hum.spe[hum_type[hum.spe$ensembl_gene_id,]$gene_biotype=="protein_coding",]
```

### Human specific expression
```{r}
hum.spe.det <- intersect(hum.spe$external_gene_name.x, rownames(dge.hum.germ.raw))
hum.spe.det.coding <- intersect(hum.spe.coding$external_gene_name.x, rownames(dge.hum.germ.raw))
```

#One to many
```{r}
hum.onetomany <- humMacMou.AllGene[humMacMou.AllGene$mmusculus_homolog_orthology_type=="ortholog_one2many" & humMacMou.AllGene$mmulatta_homolog_orthology_type=="ortholog_one2many",]

hum.onetomany.coding <- hum.onetomany[hum_type[hum.onetomany$ensembl_gene_id,]$gene_biotype=="protein_coding",]

hum.onetomany.det.coding <- hum.onetomany.coding[hum.onetomany.coding$external_gene_name.x %in% rownames(dge.hum.germ.raw),]

hum.onetomany.mac.det.coding <- intersect(hum.onetomany.coding$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw)) 

hum.onetomany.mou.det.coding <- intersect(hum.onetomany.coding$mmusculus_homolog_associated_gene_name, rownames(dge.mou.germ.raw))

hum.onetomany.mac.det <- intersect(hum.onetomany$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw)) 

hum.onetomany.mou.det <- intersect(hum.onetomany$mmusculus_homolog_associated_gene_name, rownames(dge.mou.germ.raw))

hum.onetomany.hum.det <- intersect(hum.onetomany$external_gene_name.x, rownames(dge.hum.germ.raw))

```

```{r}
write.table(unique(hum.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
write.table(unique(hum.onetomany$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.hum.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mac.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mou.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(hum.onetomany.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.hum.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.coding$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mac.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.coding$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mou.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(hum.onetomany.hum.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.hum.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.mac.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mac.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.mou.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mou.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(hum.onetomany.det.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.hum.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.mac.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mac.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.onetomany.mou.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.1m.mou.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

###. Mouse
```{r}
ensembl = useDataset("mmusculus_gene_ensembl",mart=ensembl)

attributes = c("ensembl_gene_id","external_gene_name","hsapiens_homolog_ensembl_gene","hsapiens_homolog_associated_gene_name","hsapiens_homolog_perc_id","hsapiens_homolog_perc_id_r1","hsapiens_homolog_goc_score","hsapiens_homolog_wga_coverage","hsapiens_homolog_dn","hsapiens_homolog_ds","hsapiens_homolog_orthology_confidence","hsapiens_homolog_orthology_type")

#Get all Human-Macaque gene pairs
mouHum <- getBM(attributes = attributes,mart = ensembl)
dim(mouHum)

mou_type <- getBM(attributes = c("ensembl_gene_id","external_gene_name","gene_biotype"),mart = ensembl)

###. Mouse-Macaque
attributes = c("ensembl_gene_id","external_gene_name","mmulatta_homolog_ensembl_gene","mmulatta_homolog_associated_gene_name", "mmulatta_homolog_perc_id","mmulatta_homolog_perc_id_r1","mmulatta_homolog_goc_score","mmulatta_homolog_wga_coverage","mmulatta_homolog_dn","mmulatta_homolog_ds","mmulatta_homolog_orthology_confidence","mmulatta_homolog_orthology_type")

#Get all Human-Macaque gene pairs
mouMac <- getBM(attributes = attributes,mart = ensembl)
dim(mouMac)

```

```{r}
mouHumMac.AllGene <- merge(mouHum, mouMac, by = "ensembl_gene_id")
rownames(mou_type) = mou_type$ensembl_gene_id
mou.spe <- mouHumMac.AllGene[mouHumMac.AllGene$mmulatta_homolog_ensembl_gene=="" & mouHumMac.AllGene$hsapiens_homolog_ensembl_gene=="",]
mou.spe.coding <- mou.spe[mou_type[mou.spe$ensembl_gene_id,]$gene_biotype=="protein_coding",]
```

### Mouse specific expression
```{r}
mou.spe.det <- intersect(mou.spe$external_gene_name.x, rownames(dge.mou.germ.raw))
mou.spe.det.coding <- intersect(mou.spe.coding$external_gene_name.x, rownames(dge.mou.germ.raw))
```

```{r}
write.table(unique(mou.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

#One to many
```{r}
mou.onetomany <- mouHumMac.AllGene[mouHumMac.AllGene$hsapiens_homolog_orthology_type == "ortholog_one2many" & mouHumMac.AllGene$mmulatta_homolog_orthology_type=="ortholog_one2many",]
mou.onetomany.coding <- mou.onetomany[mou_type[mou.onetomany$ensembl_gene_id,]$gene_biotype=="protein_coding",]
mou.onetomany.coding <- mou.onetomany.coding[mou.onetomany.coding$external_gene_name.x %in% rownames(dge.mou.germ.norm),]
mou.onetomany.mac.coding <- mou.onetomany.coding[mou.onetomany.coding$mmulatta_homolog_associated_gene_name %in% rownames(dge.mac.germ.norm),]$mmulatta_homolog_associated_gene_name
mou.onetomany.hum.coding <- mou.onetomany.coding[mou.onetomany.coding$hsapiens_homolog_associated_gene_name %in% rownames(dge.hum.germ.norm),]$hsapiens_homolog_associated_gene_name
```

###. Monkey
```{r}
ensembl = useDataset("mmulatta_gene_ensembl",mart=ensembl)
#listFilters(ensembl) #265 with_mmulatta_homolog
#searchAttributes(mart = ensembl, pattern = "mmulatta")

attributes = c("ensembl_gene_id","external_gene_name","hsapiens_homolog_ensembl_gene","hsapiens_homolog_associated_gene_name","hsapiens_homolog_perc_id","hsapiens_homolog_perc_id_r1","hsapiens_homolog_goc_score","hsapiens_homolog_wga_coverage","hsapiens_homolog_dn","hsapiens_homolog_ds","hsapiens_homolog_orthology_confidence","hsapiens_homolog_orthology_type")

#Get all Human-Macaque gene pairs
macHum <- getBM(attributes = attributes,mart = ensembl)
dim(macHum)

mac_type <- getBM(attributes = c("ensembl_gene_id","external_gene_name","gene_biotype"),mart = ensembl)

###. Mouse-Macaque
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name", "mmusculus_homolog_perc_id","mmusculus_homolog_perc_id_r1","mmusculus_homolog_goc_score","mmusculus_homolog_wga_coverage","mmusculus_homolog_dn","mmusculus_homolog_ds","mmusculus_homolog_orthology_confidence","mmusculus_homolog_orthology_type")

#Get all Human-Macaque gene pairs
macMou <- getBM(attributes = attributes,mart = ensembl)
dim(macMou)

```

```{r}
macHumMou.AllGene <- merge(macHum, macMou, by = "ensembl_gene_id")
rownames(mac_type) = mac_type$ensembl_gene_id
mac.spe <- macHumMou.AllGene[macHumMou.AllGene$mmusculus_homolog_ensembl_gene=="" & macHumMou.AllGene$hsapiens_homolog_ensembl_gene=="",]
mac.spe.coding <- mac.spe[mac_type[mac.spe$ensembl_gene_id,]$gene_biotype=="protein_coding",]
```

### Monkey specific expression
```{r}
mac.spe.det <- intersect(mac.spe$external_gene_name.x, rownames(dge.mac.germ.raw))
mac.spe.det.coding <- intersect(mac.spe.coding$external_gene_name.x, rownames(dge.mac.germ.raw))
```

```{r}
write.table(unique(mac.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

#####. primate-specific 
```{r}
hummac.spe <- humMacMou.AllGene[humMacMou.AllGene$mmulatta_homolog_ensembl_gene!="" & humMacMou.AllGene$mmusculus_homolog_ensembl_gene=="",]
hummac.spe.coding <- hummac.spe[hum_type[hummac.spe$ensembl_gene_id, ]$gene_biotype=="protein_coding",]
```

### primates specific expression
```{r}
hum.hummac.spe.det <- intersect(hummac.spe$external_gene_name.x, rownames(dge.hum.germ.raw))
mac.hummac.spe.det <- intersect(hummac.spe$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))

hum.hummac.spe.det.coding <- intersect(hummac.spe.coding$external_gene_name.x,rownames(dge.hum.germ.raw))
mac.hummac.spe.det.coding <-intersect(hummac.spe.coding$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))
```

```{r}
write.table(unique(hummac.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummac.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hummac.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummac.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.hummac.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummac.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.hummac.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummac.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(hummac.spe$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.hummac.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hummac.spe.coding$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.hummac.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.hummac.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.hummac.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.hummac.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.hummac.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

#####. moumac-specific
```{r}
moumac.spe <- mouHumMac.AllGene[mouHumMac.AllGene$hsapiens_homolog_ensembl_gene=="" & mouHumMac.AllGene$mmulatta_homolog_ensembl_gene!="",]
moumac.spe.coding <- moumac.spe[mou_type[moumac.spe$ensembl_gene_id, ]$gene_biotype=="protein_coding",]
```

### moumac specific expression
```{r}
mou.moumac.spe.det <- intersect(moumac.spe$external_gene_name.x, rownames(dge.mou.germ.raw))
mac.moumac.spe.det <- intersect(moumac.spe$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))

mou.moumac.spe.det.coding <- intersect(moumac.spe.coding$external_gene_name.x,rownames(dge.mou.germ.raw))
mac.moumac.spe.det.coding <-intersect(moumac.spe.coding$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))
```

```{r}
write.table(unique(moumac.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.moumac.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(moumac.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.moumac.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.moumac.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.moumac.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.moumac.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.moumac.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(moumac.spe$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.moumac.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(moumac.spe.coding$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.moumac.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.moumac.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.moumac.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mac.moumac.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mac.moumac.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

#####. hummou-specific
```{r}
hummou.spe <- humMacMou.AllGene[humMacMou.AllGene$mmulatta_homolog_ensembl_gene=="" & humMacMou.AllGene$mmusculus_homolog_ensembl_gene!="", ]
hummou.spe.coding <- hummou.spe[hum_type[hummou.spe$ensembl_gene_id, ]$gene_biotype=="protein_coding",]
```

```{r}
mou.hummou.spe.det <- intersect(hummou.spe$mmusculus_homolog_associated_gene_name,rownames(dge.mou.germ.raw))
hum.hummou.spe.det <- intersect(hummou.spe$external_gene_name.x,rownames(dge.hum.germ.raw))

mou.hummou.spe.det.coding <- intersect(hummou.spe.coding$mmusculus_homolog_associated_gene_name,rownames(dge.mou.germ.raw))
hum.hummou.spe.det.coding <-intersect(hummou.spe.coding$external_gene_name.x, rownames(dge.hum.germ.raw))
```

```{r}
write.table(unique(hummou.spe$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummou.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hummou.spe.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummou.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.hummou.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummou.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hum.hummou.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hum.hummou.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(hummou.spe$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.hummou.spe.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(hummou.spe.coding$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.hummou.spe.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.hummou.spe.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.hummou.spe.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(unique(mou.hummou.spe.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/mou.hummou.spe.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```
### 1-1-1
###. Human-Macaque
```{r}
library(biomaRt) #Ensemble Gene 94
listEnsemblArchives()
#listMarts()
ensembl <- useMart(host='http://oct2018.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL')
#ensembl = useEnsembl(biomart="ensembl",version=94)
#listDatasets(ensembl)
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
#listFilters(ensembl) #265 with_mmulatta_homolog
#searchAttributes(mart = ensembl, pattern = "mmulatta")

attributes = c("ensembl_gene_id","external_gene_name","mmulatta_homolog_ensembl_gene","mmulatta_homolog_associated_gene_name","mmulatta_homolog_perc_id","mmulatta_homolog_perc_id_r1","mmulatta_homolog_goc_score","mmulatta_homolog_wga_coverage","mmulatta_homolog_dn","mmulatta_homolog_ds","mmulatta_homolog_orthology_confidence","mmulatta_homolog_orthology_type")

#Get all Human-Macaque gene pairs
humMac <- getBM(attributes = attributes,mart = ensembl)
dim(humMac)


#Remove non-homologs
humMacOrth <- humMac[humMac$mmulatta_homolog_ensembl_gene!='',]
dim(humMacOrth)  ##23781

#Confidence
sum(humMacOrth$mmulatta_homolog_orthology_confidence == 1)
#20895  High confidence
sum(humMacOrth$mmulatta_homolog_orthology_confidence == 0)
#2976  Low confidence
```

###. Human-Mouse
```{r}
#searchAttributes(mart = ensembl, pattern = "mmusculus")
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name", "mmusculus_homolog_perc_id","mmusculus_homolog_perc_id_r1","mmusculus_homolog_goc_score","mmusculus_homolog_wga_coverage","mmusculus_homolog_dn","mmusculus_homolog_ds","mmusculus_homolog_orthology_confidence","mmusculus_homolog_orthology_type")

humMou <- getBM(attributes = attributes,mart = ensembl)
dim(humMou)

#Remove non-homologs
humMouOrth <- humMou[humMou$mmusculus_homolog_ensembl_gene!='',]
dim(humMouOrth)  ##26312

#Confidence
sum(humMouOrth$mmusculus_homolog_orthology_confidence == 1)
#17472  High confidence
sum(humMouOrth$mmusculus_homolog_orthology_confidence == 0)
#8840  Low confidence
```


###. Human-Macaque-Mouse

```{r}
humMacMouOrth <- merge(humMacOrth,humMouOrth,by = "ensembl_gene_id")
```

###. replace genes that have no symbol in macaque with human gene symbols

```{r}
emprows = which(humMacMouOrth$mmulatta_homolog_associated_gene_name == '')
humMacMouOrth$mmulatta_homolog_associated_gene_name[emprows] = humMacMouOrth$external_gene_name.x[emprows]
```

####. One-to-one-to-one genes
```{r}
humMacMouOrth.1to1 <- humMacMouOrth[humMacMouOrth$mmulatta_homolog_orthology_type=="ortholog_one2one",]
humMacMouOrth.1to1 <- humMacMouOrth.1to1[humMacMouOrth.1to1$mmusculus_homolog_orthology_type=="ortholog_one2one",]
```

```{r}
humMacMouOrth.1to1.hum.coding <- humMacMouOrth.1to1[hum_type[humMacMouOrth.1to1$ensembl_gene_id,]$gene_biotype=="protein_coding",]

humMacMouOrth.1to1.mac.coding <- humMacMouOrth.1to1[mac_type[humMacMouOrth.1to1$mmulatta_homolog_ensembl_gene,]$gene_biotype=="protein_coding",]

humMacMouOrth.1to1.mou.coding <- humMacMouOrth.1to1[mou_type[humMacMouOrth.1to1$mmusculus_homolog_ensembl_gene,]$gene_biotype=="protein_coding",]
```

```{r}
humMacMouOrth.1to1.hum.det <- intersect(humMacMouOrth.1to1$external_gene_name.x, rownames(dge.hum.germ.raw))

humMacMouOrth.1to1.mac.det <- intersect(humMacMouOrth.1to1$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))

humMacMouOrth.1to1.mou.det <- intersect(humMacMouOrth.1to1$mmusculus_homolog_associated_gene_name, rownames(dge.mou.germ.raw))

humMacMouOrth.1to1.hum.det.coding <- intersect(humMacMouOrth.1to1.hum.coding$external_gene_name.x, rownames(dge.hum.germ.raw))

humMacMouOrth.1to1.mac.det.coding <- intersect(humMacMouOrth.1to1.mac.coding$mmulatta_homolog_associated_gene_name, rownames(dge.mac.germ.raw))

humMacMouOrth.1to1.mou.det.coding <- intersect(humMacMouOrth.1to1.mou.coding$mmusculus_homolog_associated_gene_name, rownames(dge.mou.germ.raw))
```

```{r}
write.table(unique(humMacMouOrth.1to1$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.hum.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mac.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mou.db.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.hum.coding$external_gene_name.x),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.hum.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mac.coding$mmulatta_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mac.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mou.coding$mmusculus_homolog_associated_gene_name),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mou.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.hum.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.hum.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mac.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mac.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mou.det),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mou.det.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.hum.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.hum.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mac.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mac.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(unique(humMacMouOrth.1to1.mou.det.coding),file="/home/xianingz/storage/HumanMacaqueMouseCompAnaysis/interdata/orthologues_categories/hummacmou.mou.det.coding.txt",quote = FALSE, row.names = FALSE, col.names = FALSE)
```

