
#Ligand-receptor analysis for SPG cells and Somatic cells in Mouse (using ligand/receptor interaction data from Ramilowski_2015)

567 Receptors
586 Ligands
2015 Interactions

##1. Select highly variable genes from somatic and germline clusters clusters using Jun's code
```{r}
soma.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/MouseAdultST25_7SomaticCellTypes_Centroids_Qianyi.txt",header = T,sep="\t",row.names = 1))
germ.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/SPG_centroids/MouseSPGCluster_AllGene_Centroids.txt",header = T,sep="\t",row.names = 1))
##data source: https://umich.app.box.com/folder/31463886658
##First 1-Mar and 2-Mar are removed because of duplicates.
```


###. Select highly variable genes across stages.
1: SPG
2-3: prelep
4-8: spermatocytes
9-11: spermatids
12: ES
```{r}
g.mean<-apply(germ.mean,1,mean)
g.var<-apply(germ.mean,1,var)

s.mean <- apply(soma.mean,1,mean)
s.var <- apply(soma.mean,1,var)

plot(g.mean,g.var/(g.mean),cex=0.5,ylim=c(0,3),main="Germ cell")
abline(h=0.2)
abline(v=0.2)

plot(s.mean,s.var/(s.mean),cex=0.5,ylim=c(0,3),main="Somatic cell")
abline(h=0.2)
abline(v=0.2)

filter.g<-(g.mean>0.15)&( g.var/g.mean>0.1)
sum(filter.g)
filter.s<-(s.mean>0.2)&( s.var/s.mean>0.1)
sum(filter.s)
```

```{r}
somaGenes <- rownames(soma.mean)[filter.s]
germGenes <- rownames(germ.mean)[filter.g]
```

##2. Read in Ligand and Receptor data
```{r}
lrinter <- read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/Ramilowski_LigRec_Mouse.txt",header = FALSE)
ligands <- unique(lrinter$V1)
receptors <- unique(lrinter$V2)
```

##3. Extract Ligands and Receptors from Germline and Sertoli cells
```{r}
germ.ligands <- intersect(germGenes,ligands)
germ.receptors <- intersect(germGenes,receptors)
soma.ligands <- intersect(somaGenes,ligands)
soma.receptors <- intersect(somaGenes,receptors)
length(germ.ligands) #Number of ligands in germline cells
## [1] 34
length(germ.receptors) #Number of receptors in germline cells
## [1] 30
length(soma.ligands) #Number of ligands in somatic cells
## [1] 126
length(soma.receptors) #Number of receptors in somatic cells
## [1] 133
```

##4. Find ligand receptor pairs from highly variable genes
```{r}
ligrec <- function(ligands,receptors){
  lig.rec.pairs <- c()
  for(i in ligands){
    rec.db <- lrinter[lrinter[,1]==i,2]
    recs <- intersect(rec.db,receptors)
    for(j in recs){
      lig.rec.pairs <- rbind(lig.rec.pairs,c(i,j))
    }
  }
  return(lig.rec.pairs)
}
lig.rec.germ.soma <- ligrec(germ.ligands,soma.receptors)
lig.rec.soma.germ <- ligrec(soma.ligands,germ.receptors)
lig.rec.soma.soma <- ligrec(soma.ligands,soma.receptors)
lig.rec.germ.germ <- ligrec(germ.ligands,germ.receptors)
dim(lig.rec.germ.soma)[1] # Number of pairs for Ligand(Germ)-Receptor(Soma)
## [1] 56
dim(lig.rec.soma.germ)[1] # Number of pairs for Ligand(Soma)-Receptor(Germ)
## [1] 60
dim(lig.rec.soma.soma)[1] # Number of pairs for Ligand(Soma)-Receptor(Soma)
## [1] 228
dim(lig.rec.germ.germ)[1]
## [1] 12
```

##7. Use multiplication of expression values as interaction score
```{r}
mulInt <- function(inters, ligandexp, receptorexp){
  summatrix <- matrix(0, nrow = dim(ligandexp)[2], ncol = dim(receptorexp)[2])
  for(i in c(1:dim(inters)[1])){
    #c(lig,rec) <- inters[i,]
    lig <- inters[i,1]
    rec <- inters[i,2]
    summatrix <- summatrix + ligandexp[lig,] %*% t(receptorexp[rec,])
  }
  rownames(summatrix) <- c(1:dim(summatrix)[1])
  colnames(summatrix) <- c(1:dim(summatrix)[2])
  return(summatrix)
}
lig.rec.germ.soma.mul <- mulInt(lig.rec.germ.soma, germ.mean, soma.mean)
lig.rec.soma.germ.mul <- mulInt(lig.rec.soma.germ, soma.mean, germ.mean)
lig.rec.soma.soma.mul <- mulInt(lig.rec.soma.soma, soma.mean, soma.mean)
```

```{r}
library(ComplexHeatmap)
```

```{r}
soma.lab <- c("InnateLymphoid","Macrophage","Endothelial","Myoid","Leydig","Sertoli","Unknown")
colnames(lig.rec.germ.soma.mul) <- soma.lab
rownames(lig.rec.soma.germ.mul) <- soma.lab
rownames(lig.rec.soma.soma.mul) <- soma.lab
colnames(lig.rec.soma.soma.mul) <- soma.lab
Heatmap(lig.rec.germ.soma.mul, cluster_rows = FALSE, cluster_columns = FALSE)## Ligands in Germ cells and Receptors in Somatic cells
Heatmap(lig.rec.soma.germ.mul, cluster_rows = FALSE, cluster_columns = FALSE) ## Ligands in Somatic cells and receptors in Germ cells.
Heatmap(lig.rec.soma.soma.mul, cluster_rows = FALSE, cluster_columns = FALSE)## Ligands in Somatic cells and receptors in Somatic cells.
```
##8. Interaction plot
```{r}
library(plotrix)
library(shape)
```

Calculate score, rank score and select high interactions
```{r eval=FALSE}
germ.lab <- c(1:6)
soma.lab <- c("InnateLymphoid","Macrophage","Endothelial","Myoid","Leydig","Sertoli","Unknown")
lig.lab <- germ.lab
rec.lab <- soma.lab
lig.label <- "Germ"
rec.label <- "Soma"
int.plot <- function(lig.rec, lig.mean, rec.mean, lig.lab, rec.lab, lig.label, rec.label,lineup,perc,filename){
pdf(filename,width = 6*lineup[2],height = 6*lineup[1])
par(mfrow=lineup)
for(ind in c(1:dim(lig.rec[,])[1])){
perc=perc
expmat <- mulInt(matrix(lig.rec[ind,],nrow = 1),lig.mean,rec.mean)
expframe <- as.data.frame(as.table(expmat))
expframe.sort <- expframe[order(expframe$Freq,decreasing = TRUE),]
expframe.sel <- expframe.sort[c(1:floor(dim(expframe.sort)[1]*perc)),]
lig.len = dim(lig.mean)[2]
rec.len = dim(rec.mean)[2]
plot(1, type="n", xlab="", ylab="", xlim=c(0, 12), ylim=c(0, max(lig.len,rec.len)+3),axes = FALSE) ##axes2 = FALSE
lig.exp = lig.mean[lig.rec[ind,1],]
rec.exp = rec.mean[lig.rec[ind,2],]

for(i in c(1:lig.len)){
  if(lig.len >= rec.len){
    h = i
  }else{
    h = i + ceiling((rec.len-lig.len)/2)
  }
  draw.circle(3.5,h,(1/(1.5*pi))/max(lig.exp)*rev(lig.exp)[i],border = "#5a97e3",col = "#5a97e3")
  text(3,h,labels = rev(lig.lab)[i],cex=1.2, adj=c(1,0.5))
}
text(3.5,max(lig.len,rec.len)+1.2,paste0(lig.label,"\n(ligand)"),cex=1.5)
for(i in c(1:rec.len)){
  if(lig.len <= rec.len){
    h = i
  }else{
    h = i+ceiling((lig.len-rec.len)/2)
  }
  draw.circle(8.5,h,(1/(1.5*pi))/max(rec.exp)*rev(rec.exp)[i],border = "#e3a65a",col = "#e3a65a")
  text(8.8,h,label = rev(rec.lab)[i],cex=1.2, adj=c(0,0.5))
}
text(8.5,max(lig.len,rec.len)+1.2,paste0(rec.label,"\n(receptor)"),cex = 1.5)
shrink = 0.92
for(i in c(1:dim(expframe.sel)[1])){
  if(lig.len >= rec.len){
    h1 = lig.len + 1
    #h2 = lig.len + 1 - floor((lig.len - rec.len)/2)
  }else{
    #h2 = rec.len + 1
    h1 = lig.len + 1 + ceiling((rec.len - lig.len)/2)
    #h2 = rec.len + 1
    #h1 = rec.len + 1
  }
  headx = 3.3+shrink*(5-0.6)
  heady = h1-as.numeric(expframe.sel[i,1]) + shrink*(as.numeric(expframe.sel[i,1]) - sign(lig.len - rec.len)*floor(abs(lig.len - rec.len)/2) - as.numeric(expframe.sel[i,2]))
  Arrows(3.8,h1-as.numeric(expframe.sel[i,1]),headx,heady,lwd = 5/max(expframe.sel[,3])*expframe.sel[i,3])
}
text(6,max(lig.len,rec.len)+2.4,paste(lig.rec[ind,1],lig.rec[ind,2],sep = "-"),cex=2)
}
dev.off()
}
```

```{r eval=FALSE}
int.plot(lig.rec.germ.soma, lig.mean = germ.mean, rec.mean = soma.mean, lig.lab = germ.lab, rec.lab = soma.lab, lig.label = "Germ", rec.label = "Soma",lineup = c(14,4), perc=0.2,filename="~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20190402_SPG_Soma/mouse/lig.rec.germ.soma.20per.pdf")
int.plot(lig.rec.soma.germ, lig.mean = soma.mean, rec.mean = germ.mean, lig.lab = soma.lab, rec.lab = germ.lab, lig.label = "Soma", rec.label = "Germ",lineup = c(15,4), perc=0.2,filename="~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20190402_SPG_Soma/mouse/lig.rec.soma.germ.20per.pdf")
int.plot(lig.rec.soma.soma, lig.mean = soma.mean, rec.mean = soma.mean, lig.lab = soma.lab, rec.lab = soma.lab, lig.label = "Soma", rec.label = "Soma",lineup = c(57,4), perc=0.2,filename="~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20190402_SPG_Soma/mouse/lig.rec.soma.soma.20per.pdf")
int.plot(lig.rec.germ.germ, lig.mean = germ.mean, rec.mean = germ.mean, lig.lab = germ.lab, rec.lab = germ.lab, lig.label = "Germ", rec.label = "Germ",lineup = c(3,4), perc=0.2,filename="~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20190402_SPG_Soma/mouse/lig.rec.germ.germ.20per.pdf")
```